
---
params:
  folder:
    value: x
title: "Backtest - experiment"
author: "Brent Morrison"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab 
    highlight: pygments
    toc: TRUE
    toc_depth: 3
    number_sections: FALSE
    code_folding: hide
    toc_float:
      smooth_scroll: FALSE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
folder <- params$folder

# Libraries
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(cowplot)
library(slider)
library(readr)
library(jsonlite)
library(kableExtra)
library(PerformanceAnalytics)
library(DescTools)

# Set default theme
def_theme1 <- theme_minimal() +
  theme(
    legend.title = element_blank(),
    legend.position = c(0.9,0.9),
    legend.background = element_blank(),
    legend.key = element_blank(),
    plot.caption = element_text(size = 8, color = "grey55", face = 'italic'), 
    axis.title.y = element_text(size = 8, color = "darkslategrey"),
    axis.title.x = element_text(size = 8, color = "darkslategrey"),
    axis.text.y = element_text(size = 7, color = "darkslategrey"),
    axis.text.x = element_text(size = 7, color = "darkslategrey")
    )

# External arguments / parameters
json_args <- jsonlite::read_json(paste0(getwd(),"/01-scripts_02-args.json"))
```

## Backtest parameters  

Model set-up and predictor variables.
```{r backtest_parameters1}
# Arguments
args <- stack(json_args)
args <- args[, c(2,1)]
colnames(args) <- c("Attribute","Value")
kbl(args) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


## Predictor variables  

```{r backtest_parameters2}
# Get predictor variables from model embedded in workflow
model <- readRDS(paste0(getwd(),"/03-model_01-oos-pred-model"))
#predictor_vars <- model[["fit"]][["fit"]][["fit"]][["feature_names"]] # xgboost
predictor_vars <- names(model[["pre"]][["mold"]][["predictors"]])

kbl(data.frame(predictor_vars)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Predictor descriptive statistics  

```{r chunk0}
training_data_raw <- read_csv(paste0(getwd(),"/02-data_01-training.csv"))
max_date <- max(training_data_raw$date_stamp)
cor_data <- sapply(training_data_raw[training_data_raw$date_stamp == max_date , c('fwd_rtn', predictor_vars)], Winsorize, probs = c(0.05, 0.95))
chart.Correlation(cor_data, histogram=TRUE, pch=1)
```



## Predicted vs actual returns  


```{r chunk1}
vldtn_preds_raw <- read_csv(paste0(getwd(),"/02-data_04-vldtn-preds.csv"))

p1 <- ggplot(
  vldtn_preds_raw, 
  aes(y = fwd_rtn, x = .pred)
  ) + 
  geom_point(alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1) +
  labs(
    title = 'Predicted returns',
    subtitle = 'Actual vs fitted',
    y = 'Actual',
    x = 'Fitted'
  ) +
  theme(
    plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey40", hjust = 0),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  def_theme1

p2 <- vldtn_preds_raw %>% 
  mutate(residual = .pred - fwd_rtn) %>% 
  ggplot(aes(y = residual, x = .pred)) + 
    geom_point(alpha = 0.1) +
    geom_abline(intercept = 0, slope = 1) +
    labs(
      title = 'Predicted returns',
      subtitle = 'Residual vs fitted',
      y = 'Residual',
      x = 'Fitted'
    ) +
    theme(
      plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey40", hjust = 0),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    def_theme1

plot_grid(p1, p2, align = "h")
```


## Variable importance

### Shapley values via `fastshap`

```{r chunk1.2}
shap_raw <- read_csv(paste0(getwd(),"/02-data_05-shap-values.csv"))

#shap_imp <- data.frame(
#  Variable = names(shap_raw[,!names(shap_raw) %in% c("start", "end")]),
#  Importance = apply(
#    shap_raw[,!names(shap_raw) %in% c("start", "end")],
#    MARGIN = 2, FUN = function(x) sum(abs(x))
#    )
#)

shap_imp <- shap_raw %>% 
  select(-start) %>% 
  group_by(end) %>%
  summarise(across(everything(), function(x) sum(abs(x)))) %>% 
  pivot_longer(!end, names_to = "Predictors", values_to = "Importance")
             

# Plot Shap-based variable importance
ggplot(shap_imp, aes(reorder(Predictors, Importance), Importance)) +
  geom_col() +
  coord_flip() +
  xlab("") +
  ylab("mean(|Shapley value|)") +
  facet_wrap(vars(end)) +
  labs(
    title = 'Sum of the absolute value of Shapley values',
    subtitle = 'By sliding window validation set',
    x = '',
    y = ''
  ) +
  def_theme1
```

### Model specific via `vip`

```{r chunk1.3}
vip_raw <- read_csv(paste0(getwd(),"/02-data_07-var-imp.csv"))

# Plot Shap-based variable importance
ggplot(vip_raw, aes(reorder(Variable, Importance), Importance)) +
  geom_col() +
  coord_flip() +
  xlab("") +
  ylab("mean(|Shapley value|)") +
  facet_wrap(vars(end)) +
  labs(
    title = 'Model specific variable importance',
    subtitle = 'By sliding window model specification',
    x = '',
    y = ''
  ) +
  def_theme1
```


## Out of sample predictions

Text here ....

```{r chunk2}
#params$f
df <- read_csv(paste0(getwd(),"/02-data_07-oos-preds.csv"))
df <- df[,c('date_stamp','symbol','.pred','rtn_ari_1m','rtn_ari_12m')]
df <- data.frame(lapply(df, function(x) if(is.numeric(x)) round(x, 3) else x))
rmarkdown::paged_table(df)
```

## Stability-plasticity dilemma

http://www.mlfactor.com/causality.html

## Answer

Text here ....


## Performance metrics
Add percentage of correct sign 02-data_03-sp500.csv


```{r chunk4}
sp500_raw <- read_csv(paste0(getwd(),"/02-data_03-sp500.csv")) %>% 
  arrange(date_stamp)

date_filter <- min(vldtn_preds_raw$date_stamp)

preds <- vldtn_preds_raw %>% 
  group_by(date_stamp) %>% 
  mutate(fwd_rtn_dcl = ntile(.pred, 10)) %>% 
  filter(fwd_rtn_dcl == 10) %>% 
  #group_by(date_stamp, fwd_rtn_dcl) %>% 
  summarise(
    mean_fwd_rtn = mean(fwd_rtn),
    n = n()
    ) %>% 
  ungroup() %>% 
  mutate(date_stamp = add_with_rollback(date_stamp, months(1)))

# Add row to make cumsum work
preds <- bind_rows(
  data.frame(setNames(list(date_filter,0,0), names(preds))), 
  preds
  ) %>% 
  arrange(date_stamp) %>% 
  mutate(
    fcast_cum_rtn = cumprod(1 + mean_fwd_rtn)#,
    #popn_cum_rtn = cumprod(1 + mean_rtn)
    )

sp500 <- filter(sp500_raw, date_stamp >= date_filter) %>% 
  group_by(date_stamp = floor_date(date_stamp, "month")) %>% 
  mutate(date_stamp = ceiling_date(date_stamp, unit = "month") - 1) %>% 
  summarise(adjusted_close = last(adjusted_close))

sp500 <- sp500 %>% 
  mutate(
    rtn_ari_1m = replace_na((adjusted_close-lag(adjusted_close))/lag(adjusted_close), 0),
    sp500_cum_rtn = cumprod(1 + rtn_ari_1m)
    )

#ggplot(sp500, aes(x = date_stamp, y = cum_rtn)) + 
#  geom_line() #+ labs(title=folder)

ggplot() +
  geom_line(data = sp500, aes(x = date_stamp, y = sp500_cum_rtn, color = 'S&P 500')) +
  geom_line(data = preds, aes(x = date_stamp, y = fcast_cum_rtn, color = 'Model forecast')) +
  #geom_line(data = preds, aes(x = date_stamp, y = popn_cum_rtn, color = 'Model population')) +
  labs(
    x = '',
    y = '',
    title = 'Strategy and market returns'
    ) +
  scale_x_date(date_breaks = '1 years',
               date_labels = '%Y')  +
  def_theme1 + theme(legend.position = c(0.9,0.7))
  
```



```{r chunk5}
files <- list.files()
df <- data.frame(files)
rmarkdown::paged_table(df)
```