---
title: "Backtest - experiment"
author: "Brent Morrison"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab 
    highlight: pygments
    toc: TRUE
    toc_depth: 3
    number_sections: FALSE
    code_folding: hide
    toc_float:
      smooth_scroll: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)

# Libraries
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(cowplot)
library(slider)
library(readr)
library(jsonlite)
library(kableExtra)
library(PerformanceAnalytics)
library(DescTools)

# Set default theme
def_theme1 <- theme_minimal() +
  theme(
    legend.title = element_blank(),
    legend.position = c(0.9,0.9),
    legend.background = element_blank(),
    legend.key = element_blank(),
    plot.caption = element_text(size = 8, color = "grey55", face = 'italic'), 
    axis.title.y = element_text(size = 8, color = "darkslategrey"),
    axis.title.x = element_text(size = 8, color = "darkslategrey"),
    axis.text.y = element_text(size = 7, color = "darkslategrey"),
    axis.text.x = element_text(size = 7, color = "darkslategrey")
    )

# External arguments / parameters
json_args      <- jsonlite::read_json(paste0(getwd(),"/01-scripts_02-args.json"))
preds          <- json_args$predictors
hyper_params   <- as.logical(json_args$hyper_params)
```

<br>

## Overview  
DESCRIPTION TO GO HERE.

## Backtest parameters  

Model set-up and predictor variables.
```{r backtest_parameters1}
# Arguments
args <- stack(json_args)
args <- args[, c(2,1)]
colnames(args) <- c("Attribute","Value")
kbl(args) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


## Predictor variables  

```{r backtest_parameters2}
# Get predictor variables from model embedded in workflow
model <- readRDS(paste0(getwd(),"/03-model_01-oos-pred-model"))
#predictor_vars <- model[["fit"]][["fit"]][["fit"]][["feature_names"]] # xgboost
predictor_vars <- names(model[["pre"]][["mold"]][["predictors"]])

kbl(data.frame(predictor_vars)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


### Data

```{r chunk0.0}
training_data_raw <- read_csv(paste0(getwd(),"/02-data_01-training.csv"))

training_data_raw %>% 
  group_by(date_stamp) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = date_stamp, y = n)) + 
  geom_line() +
  labs(
      x = '',
      y = '',
      title = 'Training data',
      subtitle = 'Count of stocks by date'
      ) +
  def_theme1
```


## Predictor vs returns  

```{r chunk0}
training_data_raw <- read_csv(paste0(getwd(),"/02-data_01-training.csv"))
training_data_corr <- training_data_raw %>% 
  select(date_stamp, symbol, fwd_rtn, unlist(preds)) %>% 
  pivot_longer(cols = unlist(preds), names_to = "predictor", values_to = "pred_value") %>% 
  group_by(date_stamp, predictor) %>% 
  mutate(pred_value_wins = Winsorize(pred_value, probs = c(0.025, 0.975))) %>% 
  ungroup() 
training_data_corr %>% 
  ggplot(
  aes(y = fwd_rtn, x = pred_value_wins)
  ) + 
  geom_point(alpha = 0.01) +
  #geom_abline(intercept = 0, slope = 1) +
  labs(
    title = 'Returns and predictor variables',
    subtitle = 'Data winsorised at 2.5% and 97.5% level',
    y = 'Forward returns',
    x = 'Predictor value'
  ) +
  facet_wrap(vars(predictor), scales = "free") +
  def_theme1
```





## Predicted vs actual returns  


```{r chunk1}
if (file.exists(paste0(getwd(),"/02-data_04-vldtn-preds.csv"))){

  vldtn_preds_raw <- read_csv(paste0(getwd(),"/02-data_04-vldtn-preds.csv"))
  
  p1 <- ggplot(
    vldtn_preds_raw, 
    aes(y = fwd_rtn, x = .pred)
    ) + 
    geom_point(alpha = 0.1) +
    geom_abline(intercept = 0, slope = 1) +
    labs(
      title = 'Predicted returns',
      subtitle = 'Actual vs fitted',
      y = 'Actual',
      x = 'Fitted'
    ) +
    theme(
      plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey40", hjust = 0),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    def_theme1
  
  p2 <- vldtn_preds_raw %>% 
    mutate(residual = .pred - fwd_rtn) %>% 
    ggplot(aes(y = residual, x = .pred)) + 
      geom_point(alpha = 0.1) +
      geom_abline(intercept = 0, slope = 1) +
      labs(
        title = 'Predicted returns',
        subtitle = 'Residual vs fitted',
        y = 'Residual',
        x = 'Fitted'
      ) +
      theme(
        plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey40", hjust = 0),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()
      ) +
      def_theme1
  
  plot_grid(p1, p2, align = "h")
  
} else {
  
  print('NA')
  
}
```


## Variable importance

### Shapley values via `fastshap`

```{r chunk1.2}
if (file.exists(paste0(getwd(),'/02-data_05-shap-values.csv'))) {
  
  shap_raw <- read_csv(paste0(getwd(),"/02-data_05-shap-values.csv"))
  
  #shap_imp <- data.frame(
  #  Variable = names(shap_raw[,!names(shap_raw) %in% c("start", "end")]),
  #  Importance = apply(
  #    shap_raw[,!names(shap_raw) %in% c("start", "end")],
  #    MARGIN = 2, FUN = function(x) sum(abs(x))
  #    )
  #)
  
  shap_imp <- shap_raw %>% 
    select(-start) %>% 
    group_by(end) %>%
    summarise(across(everything(), function(x) sum(abs(x)))) %>% 
    pivot_longer(!end, names_to = "Predictors", values_to = "Importance")
               
  
  # Plot Shap-based variable importance
  ggplot(shap_imp, aes(reorder(Predictors, Importance), Importance)) +
    geom_col() +
    coord_flip() +
    xlab("") +
    ylab("mean(|Shapley value|)") +
    facet_wrap(vars(end)) +
    labs(
      title = 'Sum of the absolute value of Shapley values',
      subtitle = 'Grouped by sliding window validation set',
      x = '',
      y = ''
    ) +
    def_theme1
  
} else {
  
  print('NA')
  
}
```

### Model specific via `vip`

```{r chunk1.3}
if (file.exists(paste0(getwd(),"/02-data_07-var-imp.csv"))) {

  vip_raw <- read_csv(paste0(getwd(),"/02-data_07-var-imp.csv"))
  
  # Plot Shap-based variable importance
  ggplot(vip_raw, aes(reorder(Variable, Importance), Importance)) +
    geom_col() +
    coord_flip() +
    xlab("") +
    ylab("mean(|Shapley value|)") +
    facet_wrap(vars(end)) +
    labs(
      title = 'Model specific variable importance',
      subtitle = 'By sliding window model specification',
      x = '',
      y = ''
    ) +
    def_theme1
  
} else {
  
  print('NA')
  
}
```

### Hyper-parameter stability 

If a model captures the true data generating process the hyper-parameters should be relatively constant.  

```{r chunk1.4, eval=hyper_params}
if (file.exists(paste0(getwd(),"/02-data_06-hyper-parms.csv"))) {

  hyper_parms_raw <- read_csv(paste0(getwd(),"/02-data_06-hyper-parms.csv"))
  
  # Plot Shap-based variable importance
  hyper_parms_raw %>% 
    #pivot_longer(!date_stamp, names_to = "predictor", values_to = "coefficient") %>% 
    group_by(.config, hidden_units, penalty, end) %>% 
    summarise(mae = mean(.estimate)) %>% 
  ggplot(aes(x = end, y = mae, group = .config, color = .config)) + #, linetype = .config
    geom_line() +
    labs(
      title = "Mean absolute error by hyperparameter setting and sliding window"
      #subtitle = "",
      #caption = ""
      ) +
    def_theme1 +
    theme(legend.position = "none") +
    scale_color_grey(start = 0.9, end = 0.1) +
    scale_x_date(date_breaks = "year", date_labels = "%Y")
  
} else {
  
  print('NA')
  
}
```

## Out of sample predictions

Text here ....

```{r chunk2}
#params$f
df <- read_csv(paste0(getwd(),"/02-data_07-oos-preds.csv"))
df <- df[,c('date_stamp','symbol','.pred','rtn_ari_1m','rtn_ari_12m')]
df <- data.frame(lapply(df, function(x) if(is.numeric(x)) round(x, 3) else x))
rmarkdown::paged_table(df)
```

## Stability-plasticity dilemma

http://www.mlfactor.com/causality.html#nonstat





## Performance metrics
Add percentage of correct sign 02-data_03-sp500.csv


```{r chunk4}
sp500_raw <- read_csv(paste0(getwd(),"/02-data_03-sp500.csv")) %>% 
  arrange(date_stamp)

date_filter <- min(vldtn_preds_raw$date_stamp)

preds <- vldtn_preds_raw %>% 
  group_by(date_stamp) %>% 
  mutate(fwd_rtn_dcl = ntile(.pred, 10)) %>% 
  filter(fwd_rtn_dcl == 10) %>% 
  #group_by(date_stamp, fwd_rtn_dcl) %>% 
  summarise(
    mean_fwd_rtn = mean(fwd_rtn),
    n = n()
    ) %>% 
  ungroup() %>% 
  mutate(date_stamp = add_with_rollback(date_stamp, months(1)))

# Add row to make cumsum work
preds <- bind_rows(
  data.frame(setNames(list(date_filter,0,0), names(preds))), 
  preds
  ) %>% 
  arrange(date_stamp) %>% 
  mutate(
    fcast_cum_rtn = cumprod(1 + mean_fwd_rtn)#,
    #popn_cum_rtn = cumprod(1 + mean_rtn)
    )

sp500 <- filter(sp500_raw, date_stamp >= date_filter) %>% 
  group_by(date_stamp = floor_date(date_stamp, "month")) %>% 
  mutate(date_stamp = ceiling_date(date_stamp, unit = "month") - 1) %>% 
  summarise(adjusted_close = last(adjusted_close))

sp500 <- sp500 %>% 
  mutate(
    rtn_ari_1m = replace_na((adjusted_close-lag(adjusted_close))/lag(adjusted_close), 0),
    sp500_cum_rtn = cumprod(1 + rtn_ari_1m)
    )

#ggplot(sp500, aes(x = date_stamp, y = cum_rtn)) + 
#  geom_line() #+ labs(title=folder)

ggplot() +
  geom_line(data = sp500, aes(x = date_stamp, y = sp500_cum_rtn, color = 'S&P 500')) +
  geom_line(data = preds, aes(x = date_stamp, y = fcast_cum_rtn, color = 'Model forecast')) +
  #geom_line(data = preds, aes(x = date_stamp, y = popn_cum_rtn, color = 'Model population')) +
  labs(
    x = '',
    y = '',
    title = 'Strategy and market returns'
    ) +
  scale_x_date(date_breaks = '1 years',
               date_labels = '%Y')  +
  def_theme1 + theme(legend.position = c(0.9,0.7))
  
```



```{r chunk5}
files <- list.files()
df <- data.frame(files)
rmarkdown::paged_table(df)
```